---
title: "BayesBinMix Census Data"
author: "Carmen B. Rodriguez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
library(tidycensus)
library(tidyverse)
library(Hmisc)
library(reshape2)
library(ggplot2)
library(psych)
library(table1)
library(Hmisc)


```

## Load census data
```{r}
censusdata_bin <- readRDS("~/BayesBinMix_Project/censusdata_bin.rds")
censusdata_bin <- readRDS("/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/censusdata_bin.rds")

names(censusdata_bin) <- c("acs5_2010_bin","acs5_2015_bin","acs5_2019_bin")

```


## Bayes Bernoulli Mixture Model

```{r}
#Load required packages 
library(BayesBinMix)
library(foreach)
library(label.switching)
library(doParallel)
library(coda)

```


### Start with 5-year ACS (2010)

For first iterations we'll set $K_{max} = 10$. The "true" number of clusters is set to $K = 10$ and the dimensionality of the multivariate distribution is equal to $d = 18$.

The following prior assumptions are imposed

$$K \sim \text{Truncated Poisson}(\lambda = 1)$$

$$p|K \sim \text{Dirichlet}(1,..., 1_K)$$
We set $\gamma_1=...= \gamma_K =\gamma > 0$ so that prior assumptions do not impose any particular information that separates the mixture components between them.



$$\theta_{kj} \sim \text{Beta}(\alpha=1, \beta=1)$$


```{r}
#As in the paper, we will run 4 parallel chains with the following temperatures: (1, 0.8, 0.6, 0.4).
#***the first chain should correspond to the actual posterior distribution, so its temperature equals to 1.

Kmax<-10 #upper bound on the number of cluster
nChains <- 4
heats <- seq(1, 0.4, length = nChains)

X1<- as.matrix(censusdata_bin$acs5_2010_bin)

##Poisson- Uniform --
acs10_1<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X1, outPrefix = "acs10_res_poiunif_1000", 
                ClusterPrior = "poisson",
                m = 1000) 

##Uniform-Uniform
acs10_1a<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X1, outPrefix = "acs10_res_uniuni_1000", 
                ClusterPrior = "uniform",
                m = 1000) 





#Change prior on theta to alpha=beta = 0.5
##Poisson-Jeffreys
acs10_2<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X1, outPrefix = "acs10_res_poijeff_1000", 
                ClusterPrior = "poisson", alpha = 0.5, beta = 0.5,
                m = 1000) 


##Uniform-Jeffreys
acs10_2a<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X1, outPrefix = "acs10_res_unijeff_1000", 
                ClusterPrior = "uniform",alpha = 0.5, beta = 0.5,
                m = 1000)


#Try 8 chains and Kmax = 20 --  SEND THIS OUTPUT TO STEPHANIE
nChains8 <- 8
heats8 <- seq(1, 0.6, length = nChains8)

acs10_3<-coupledMetropolis(Kmax=20, nChains = nChains8, heats = heats8,
                binaryData = X1, outPrefix = "acs10_k20_poiunif_1000", 
                ClusterPrior = "poisson",
                m = 1000) 

check<-readRDS(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_Kmax20.rds")


### Explore output from these models
#rawMCMC_poiuni<- read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/rawMCMC.mapK.10.txt", header = T)

#Visualize the output using coda package
#Trace plots and histogram of theta_kj, p_k for 1st component
summary(poires2$parameters.ecr.mcmc)
mat1 <- matrix(c(1:4), byrow = TRUE, ncol = 2)
layout(mat1, widths = rep(c(2, 1), 2), heights = rep(1, 4))
mcmcSubset <- poires2$parameters.ecr.mcmc[ , c("theta.1.1", "p.1")]
plot(mcmcSubset, auto.layout = FALSE, ask = FALSE, col = "gray40")


#Plots like Figure 4 in the paper

#(1) Illustrate the sampled values of K per chain according to the Poisson and uniform prior distribution using data from K.allChains

poiKallchains<-read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/K.allChains.txt", header = T)

  
poiKallchains$m <-as.numeric(row.names(poiKallchains))
poiKallchains1<-poiKallchains %>% pivot_longer(cols = 1:4, names_to ="Chain", values_to = "K")
  

poijKallchains<-read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poijeff_1000/K.allChains.txt", header = T)
poijKallchains$m <-as.numeric(row.names(poijKallchains))
poijKallchains1<-poijKallchains %>% pivot_longer(cols = 1:4, names_to ="Chain", values_to = "K")



#this plot shows the generated values of K per heated chain-- note chain 1 is the original (^1)
poiKallchains1 %>% ggplot(aes(x = m, y = K, col = as.factor(Chain))) + 
  geom_line() + 
  ggtitle("MC^3 Sampler Poi-Unif") + 
  xlab("mcmc cycle")
poijKallchains1 %>% ggplot(aes(x = m, y = K, col = as.factor(Chain))) +
 geom_line() + 
  ggtitle("MC^3 Sampler Poi- Jeff") +
  xlab("mcmc cycle")



#(2) Raw output of p1,...,pK  and theta_kj conditional on K=10 & K=8 (NOT REORDERED!)-- rawMCMC.mapK.10.txt

poiraw1<-read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/rawMCMC.mapK.10.txt", header = T)
poiraw2<-read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poijeff_1000/rawMCMC.mapK.8.txt", header = T)

poiraw1a<-poiraw1[,paste0("p.", seq(1,10), "")]
colnames(poiraw1a)<-paste("Component", seq(1,10), "")
poiraw1a$iteration<-as.numeric(row.names(poiraw1a))
poiraw1a<-poiraw1a%>% pivot_longer(cols = 1:10, names_to = "Component", values_to = "raw.mix.prop")
poiraw1a %>% ggplot(aes(x=iteration, y = raw.mix.prop, col = as.factor(Component))) + geom_point()



poiraw2a<-poiraw2[,paste0("p.", seq(1,8), "")]
colnames(poiraw2a)<-paste("Component", seq(1,8), "")
poiraw2a$iteration<-as.numeric(row.names(poiraw2a))
poiraw2a<-poiraw2a%>% pivot_longer(cols = 1:8, names_to = "Component", values_to = "raw.mix.prop")
poiraw2a %>% ggplot(aes(x=iteration, y = raw.mix.prop, col = as.factor(Component))) + geom_point()


#(3) Reordered by ECR algorithm
poi1ireordered<-read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/reorderedMCMC-ECR.mapK.10.txt", header = T)

poi1ireorderedp<-poi1ireordered[,paste0("p.", seq(1,10), "")]
colnames(poi1ireorderedp)<-paste("Component", seq(1,10), "")
poi1ireorderedp$iteration<-as.numeric(row.names(poi1ireorderedp))
poi1ireorderedp<-poi1ireorderedp %>% pivot_longer(cols = 1:10, names_to = "Component", values_to = "reordered.mix.prop")

poi1ireorderedp %>% ggplot(aes(x=iteration, y = reordered.mix.prop, col = as.factor(Component))) + 
  geom_point() +
  ggtitle("Reordered sample (mixing weights) according to ECR Algorith (m=1000; 4 chains)") +
  xlab("Iteration")+
  ylab("reordered weight") + theme(legend.title=element_blank()) 



#(4) Reordered by STEPHENS algorithm
poi1ireordered_st<-read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/reorderedMCMC-STEPHENS.mapK.10.txt", header = T)

poi1ireorderedp_st<-poi1ireordered_st[,paste0("p.", seq(1,10), "")]
colnames(poi1ireorderedp_st)<-paste("Component", seq(1,10), "")
poi1ireorderedp_st$iteration<-as.numeric(row.names(poi1ireorderedp_st))
poi1ireorderedp_st<-poi1ireorderedp_st %>% pivot_longer(cols = 1:10, names_to = "Component", values_to = "reordered.mix.prop")

poi1ireorderedp_st %>% ggplot(aes(x=iteration, y = reordered.mix.prop, col = as.factor(Component))) + 
  geom_point() +
  ggtitle("Reordered sample according to STEPHENS Algorith(m=1000; 4 chains)") +
  xlab("Iteration")+
  ylab("reordered weight") + theme(legend.title=element_blank()) 



#(5) Estimated posterior distribution of the number of clusters when using different parameters (alpha,beta) on the beta prior for theta_kj. Both choices of the prior on the number of clusters are considered: truncated Poisson and uniform--- NEED. allocations.ecr.mcmc

pvarK_poiuni<- read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/p.varK.txt", header = T)

pvarK_poijeff<- read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poijeff_1000/p.varK.txt", header = T)

pvarK_uniuni<- read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_uniuni_1000/p.varK.txt", header = T)

pvarK_unijeff<- read.table(file = "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_unijeff_1000/p.varK.txt", header = T)



#(6) classificationProbabilities.mapK.KVALUE.csv 
#the reordered classification probabilities per observation after reordering the most probable number of clusters with the ECR algorithm.
poiuni_classprob<- read.csv(file= "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/classificationProbabilities.mapK.10.csv")
poiuni_classprob$CT<- as.numeric(row.names(poiuni_classprob))
colnames(poiuni_classprob)<-c(paste("Component", seq(1,10)), "CT")


poiuni_classprob_long<- poiuni_classprob[1:50,] %>% pivot_longer(cols = 1:10, names_to = "Component", values_to = "classprob")

library(RColorBrewer)
poiuni_classprob_long %>% ggplot(aes(x = CT, y = classprob, fill = Component)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette="Paired")







#(7) z-ECR.mapK.KVALUE.txt 
#the reordered simulated latent allocations which corresponds to the most probable model, reordered according to the ECR algorithm.
poiuni_allocation<- read.table(file= "/Users/carmenrodriguez/Desktop/Research Projects/BayesBinMix/BayesBinMix_FASRC/acs10_res_poiunif_1000/reorderedSingleBestClusterings.mapK.10.txt", header = T)
Zi_ECR <-poiuni_allocation$ECR




#Data frame with cluster prob and Z
dat<-cbind(poiuni_classprob, Zi_ECR)



##Want to qualitatively label clusters based on the profile of census tracts


```

### ACS 2010- 2015
```{r}
## Second dataset
X2<- as.matrix(censusdata_bin$acs5_2015_bin)

##Poisson- Uniform
acs15_1<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X2, outPrefix = "acs15_res_poiunif_1000", 
                ClusterPrior = "poisson",
                m = 1000) 
##Uniform-Uniform
acs15_1a<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X2, outPrefix = "acs15_res_uniuni_1000", 
                ClusterPrior = "uniform",
                m = 1000) 

#Change prior on thera to alpha=beta = 0.5
##Poisson-Jeffreys
acs15_2<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X2, outPrefix = "acs15_res_poijeff_1000", 
                ClusterPrior = "poisson", alpha = 0.5, beta = 0.5,
                m = 1000) 

##Uniform-Jeffreys
acs15_2a<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X2, outPrefix = "acs15_res_unijeff_1000", 
                ClusterPrior = "uniform",alpha = 0.5, beta = 0.5,
                m = 1000)
```


```{r}
## Third dataset
X3<- as.matrix(censusdata_bin$acs5_2019_bin)
##Poisson- Uniform
acs19_1<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X3, outPrefix = "acs19_res_poiunif_1000", 
                ClusterPrior = "poisson",
                m = 1000) 
##Uniform-Uniform
acs19_1a<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X3, outPrefix = "acs19_res_uniuni_1000", 
                ClusterPrior = "uniform",
                m = 1000) 

#Change prior on thera to alpha=beta = 0.5
##Poisson-Jeffreys
acs19_2<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X3, outPrefix = "acs19_res_poijeff_1000", 
                ClusterPrior = "poisson", alpha = 0.5, beta = 0.5,
                m = 1000) 


##Uniform-Jeffreys
acs19_2a<-coupledMetropolis(Kmax=Kmax, nChains = nChains, heats = heats,
                binaryData = X3, outPrefix = "acs19_res_unijeff_1000", 
                ClusterPrior = "uniform",alpha = 0.5, beta = 0.5,
                m = 1000) 
```

```{r}
#m is the number of mcmc cycles -- each cycle consists of 10 iterations and so 10000 iterations total for now-- we do not set a burn in and do further exploration to decide

#EXPLORE OUTPUT
###-------------- TRACE PLOTS---------------

#Visualize the output using coda package
#Trace plots and histogram of theta_kj, p_k for 1st component
# mat <- matrix(c(1:4), byrow = TRUE, ncol = 2)
# layout(mat, widths = rep(c(2, 1), 2), heights = rep(1, 4))
# mcmcSubset <- acs10_1$parameters.ecr.mcmc[ , c("theta.1.1", "p.1")]
# plot(mcmcSubset, auto.layout = FALSE, ask = FALSE, col = "blue")
# 
# acs10_1$classificationProbabilities.ecr
# acs10_1$parameters.ecr.mcmc

```


  