---
title: "BayesBinMix Census Data Results Summarized"
author: "Carmen B. Rodriguez"
format: 
  html:
    self-contained: TRUE
editor: visual
---

```{r}
#Load packages
library(tidyverse)
library(Hmisc)
library(reshape2)
library(ggplot2)
library(psych)
library(table1)
library(tableone)
library(Hmisc)
library(BayesBinMix)
library(foreach)
library(label.switching)
library(doParallel)
library(coda)
library(DescTools)
library(MCMCvis)
```

```{r}
#Load census data
censusdata_bin <- readRDS("/homes6/carmen/Other projects/ecbayesbinmix/censusdata_bin.rds")
names(censusdata_bin) <- c("acs5_2010_bin","acs5_2015_bin","acs5_2019_bin")

```

## Start with 5-year ACS (2006-2010)

For first iterations we'll set $K_{max} = 10$. The "true" number of clusters is set to $K = 10$ and the dimensionality of the multivariate distribution is equal to $d = 14$.

The following prior assumptions are imposed

$$K \sim \text{Truncated Poisson}(\lambda = 1)$$

$$p|K \sim \text{Dirichlet}(1,..., 1_K)$$ We set $\gamma_1=...= \gamma_K =\gamma > 0$ so that prior assumptions do not impose any particular information that separates the mixture components between them.

$$\theta_{kj} \sim \text{Beta}(\alpha=1, \beta=1)$$

## Run models

We will run models with different values of $K_{max}$ . That is, we will look at output where the upper bound on $K$ is `K= c(5,10,15)` .

```{r}
#Load model results--will be stored in GitHub
res_all <- readRDS("/homes6/carmen/Other projects/ecbayesbinmix/ACS10_results.rds")

```

### Explore output

##### (1) Trace plots-selecting a cluster at random for examination

```{r}
set.seed(2020) #year-month-day
#-------Explore output---------
mapK<-Mode(res$K.mcmc)[1]
n_ses<-ncol(X10)

## (1) Trace plots for convergence exploration -- select a cluster at random
clusterran<-sample(1:mapK, size = 1)
clustersel<-c(paste0("theta", sep=".", clusterran, sep=".",1:n_ses), paste("p", clusterran, sep = "."))
params_est<-res$parameters.ecr.mcmc

#Filename is for most probable K,not the cluster being plotted
MCMCtrace(params_est[,clustersel], filename = paste("MCMCtrace_mapK", mapK, sep = "_"), wd = "/homes6/carmen/Other projects/ecbayesbinmix/Figures")
```

##### (2) Illustrate the sampled values of K per chain according to the Poisson and uniform prior distribution using data from K.allChains--- This we can do last if necessary, otherwise we don't run it

```{r}
#| eval: false
poiKallchains<-read.table(file = "/homes6/carmen/Other projects/acs10_res_poiunif_demo/K.allChains.txt",header = T)

poiKallchains$m <-as.numeric(row.names(poiKallchains))
poiKallchains1<-poiKallchains %>% pivot_longer(cols = 1:4, names_to ="Chain", values_to = "K")
#this plot shows the generated values of K per heated chain-- note chain 1 is the original (^1)
poiKallchains1 %>% ggplot(aes(x = m, y = K, col = as.factor(Chain))) + 
  geom_line() + 
  ggtitle("MC^3 Sampler Poi-Unif") + 
  xlab("mcmc cycle")

```

### **Plots** (Patterns)

```{r}
#| fig-cap: "Success Probabilities of Neighborhood SES variables for each cluster"

#Import functions from plots.R
source("plots.R")
library(wesanderson)
library(viridis)

sesvars<-c("Household:Two People Per Room","Household: Lack of complete Plumbing", "Household:Median Income", ">= Bacherlor's Degree ", "Below Poverty Line", "< High School",
           "Female Household", ">= High School","Household:No Vehicle","Household: Owner", "Household:Renter", "Unemployment","White Collar Occupation", "SNAP Benefits")



# Other color options
viridis_palette <- viridis(14)
custom_colors <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", 
                   "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf",
                   "#ff0000", "#00ff00", "#0000ff", "#ffff00")

# Generate 14 pastel colors
pastel_palette <- generate_pastel_colors(14)

plot_thetakj(reslst = res, mapK = 9, sesvars= sesvars, color_palette = pastel_palette)

```
