---
title: "BayesBinMix Census Data (New models)"
author: "Carmen B. Rodriguez"
format: html
editor: visual
---

```{r}
#| warning: false
#| output: false
#Load packages
library(tidyverse)
library(Hmisc)
library(reshape2)
library(ggplot2)
library(psych)
library(table1)
library(tableone)
library(Hmisc)
library(BayesBinMix)
library(foreach)
library(label.switching)
library(doParallel)
library(coda)
library(DescTools)
library(MCMCvis)
library(knitr)
library(forcats)
library(gmodels)
```

The Multivariate Bernoulli Mixture Model

The dimensionality of the multivariate distribution is equal to $d = 14$.

The following prior assumptions are imposed

$$K \sim \text{Truncated Poisson}(\lambda = 1)$$

$$p|K \sim \text{Dirichlet}(\gamma_1,..., \gamma_K)$$ We set $\gamma_1=...= \gamma_K =\gamma = 1/50$. We do this so that the prior distribution on the weight distribution of a mixture with $$K$$ components is chosen in such a way that a priori the number of clusters in the data is random and is allowed to be smaller than $$K$$ with high probability.<https://link.springer.com/article/10.1007/s11634-018-0329-y>

$$\theta_{kj} \sim \text{Beta}(\alpha=1, \beta=1)$$

```{r}
#Load census data and results from latest model
censusdata_bin <- readRDS("./censusdata_bin_rownames.rds")
names(censusdata_bin) <- c("acs5_2010_bin","acs5_2015_bin","acs5_2019_bin")

res_all <- readRDS("./")




```
