---
title: "Maps"
author: "Carmen Rodriguez Cabrera"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(cache = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
#--------------Load the tidycensus package into your R session-----------------
library(tidycensus)
library(tidyverse)
library(Hmisc)
library(reshape2)
library(ggplot2)
library(sf)
library(tigris)
library(crsuggest)

# Define your Census API key and set it with census_api_key()

#readRenviron("~/.Renviron") #run this to use  API key

#Check API key
#Sys.getenv("CENSUS_API_KEY")
```


##  MAPS - Race/ Ethnicity Distribution across CTs
https://walker-data.com/tidycensus/articles/spatial-data.html

**tidycensus** can return simple feature geometry for geographic units along with 
variables from the decennial US Census or American Community survey.  By setting _geometry = TRUE_ in a tidycensus function call, tidycensus will use the tigris package to retrieve the corresponding geographic dataset from the US 
Census Bureau and pre-merge it with the tabular data obtained from the Census API.


###---- Overall 

```{r echo=FALSE,cache =TRUE}
#----------------------- MAPS-------------------------------
#https://walker-data.com/tidycensus/articles/spatial-data.html
## tidycensus can return simple feature geometry for geographic units along with 
# variables from the decennial US Census or American Community survey. 
# By setting geometry = TRUE in a tidycensus function call, tidycensus will use 
# the tigris package to retrieve the corresponding geographic dataset from the US 
# Census Bureau and pre-merge it with the tabular data obtained from the Census API.

options(tigris_use_cache = TRUE)

#Overall race breakdown
#https://www.bostonglobe.com/2021/08/12/metro/new-census-data-see-how-each-mass-town-has-changed/
racevars<- c(`Non-Hispanic White` = "B03002_003",
             `Non-Hispanic Black` = "B03002_004",
              `Non-Hispanic AIAN` = "B03002_005",
              `Non-Hispanic Asian` = "B03002_006",
              `Non-Hispanic NHPI`="B03002_007",
             `Hispanic/Latino` = "B03002_012")
race<-get_acs(state = "MA", geography = "tract", 
                  variables = racevars, year = 2019, geometry = TRUE, 
              summary_var ="B03002_001")
race<-race %>% mutate(Percentage = 100 * (estimate / summary_est))
head(race)
```


```{r eval=FALSE, cache=TRUE, include=FALSE}
# Maps showing the distribution of race and ethnicity across census tracts in MA

#Look at distribution across census tracts
# race %>% ggplot( aes(x = Percentage)) + geom_histogram( fill = "white", colour = "black") + facet_wrap( ~ variable) + ylab("MA Census Tract Count")

#pdf("/Users/carmenrodriguez/Desktop/Spring2022/BST300/Spring 2022/R files/Figures/Race_EthnicityHistograms_MAPS.pdf")
race %>% filter(variable == "Non-Hispanic White") %>% ggplot( aes(x = Percentage)) + 
  geom_histogram( fill = "white", colour = "black") + ylab("Freqcuency") + 
  ggtitle("Distribution of Non-Hispanic Whites  across MA Census tracts")

race %>% filter(variable == "Non-Hispanic Black") %>% ggplot( aes(x = Percentage)) + 
  geom_histogram( fill = "white", colour = "black") + ylab("Freqcuency") +
  ggtitle("Distribution of Non-Hispanic Black  across MA Census tracts")

race %>% filter(variable == "Non-Hispanic Asian") %>% ggplot( aes(x = Percentage)) + 
  geom_histogram( fill = "white", colour = "black") + ylab("Freqcuency") +
  ggtitle("Distribution of Non-Hispanic Asian  across MA Census tracts")

race %>% filter(variable == "Non-Hispanic AIAN")%>% ggplot( aes(x = Percentage)) + 
  geom_histogram( fill = "white", colour = "black") + ylab("Freqcuency") + 
  ggtitle("Distribution of Non-Hispanic AIAN  across MA Census tracts")

race %>% filter(variable == "Non-Hispanic NHPI") %>% ggplot( aes(x = Percentage)) + 
  geom_histogram( fill = "white", colour = "black") + ylab("Freqcuency") + 
  ggtitle("Distribution of Non-Hispanic NHPI across MA Census tracts")

race %>% filter(variable == "Hispanic/Latino") %>% ggplot( aes(x = Percentage)) + 
  geom_histogram( fill = "white", colour = "black") + ylab("Freqcuency") +
  ggtitle("Distribution of Hispanic or Latinos  across MA Census tracts")


race %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = " Race/Ethnicity by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

#dev.off()

```


```{r echo=FALSE, message=FALSE, warning=FALSE, cache =TRUE}
# Check coordinates reference system before plotting maps--which specifies not only how data coordinates should be mapped to a model of the 
MA_tracts <- tracts("MA", cb = TRUE)
st_crs(MA_tracts)
ma_crs<-crsuggest::suggest_crs(MA_tracts) #6491	NAD83(2011) / Massachusetts Mainland Zone meters

# coord_sf(crs = 6491  Help with projection/CRS
# coords_sf(datum = NA) Remove gridlines, i.e graticules

##Colors
#https://colorbrewer2.org/#type=sequential&scheme=YlOrRd&n=3
#https://ggplot2.tidyverse.org/reference/scale_viridis.html


race %>% filter(variable %in% c("Non-Hispanic White","Non-Hispanic Black", "Non-Hispanic Asian","Hispanic/Latino" )) %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = " Race/Ethnicity by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()


# race %>% ggplot(aes(fill = Percentage)) + 
#   facet_wrap(~ variable) +
#   geom_sf(color = NA) + 
#   coord_sf(crs = 6491, datum =NA) +
#   scale_fill_distiller(palette = "YlOrRd", direction = -1) +
#   labs(title = " Race/Ethnicity by Census Tracts in MA",
#        caption = "Data source: 2019 5-year ACS",
#        fill = "Percent") 

```

###---- Hispanic/Latino Heritage

```{r}
# OVERALL GROUPS
hispanic_overall<-c(
`Mexican`	= "B03001_004",
`Puerto Rican`=	"B03001_005",
`Cuban`	= "B03001_006",
`Dominican` =	"B03001_007",
`Central American`=	"B03001_008",
`South American`=	"B03001_016",
`Other Hispanic or Latino`=	"B03001_027")

options(tigris_use_cache = TRUE)
hispanic_overall<-get_acs(state = "MA", geography = "tract", 
                  variables = hispanic_overall, year = 2019, geometry = TRUE, 
                  summary_var = "B03001_003")

hispanic_overall<-hispanic_overall%>% mutate(Percentage = 100 * (estimate / summary_est))
head(hispanic_overall)

#pdf("/Users/carmenrodriguez/Desktop/Spring2022/BST300/Spring 2022/R files/Figures/hispanicAll.pdf")
hispanic_overall %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = " Hispanic or Latino: All groups by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()
#dev.off()
```


```{r eval=FALSE, include=FALSE}
#Central America
hispanic_CA<-c(`Costa Rican`=	"B03001_009",
`Guatemalan`=	"B03001_010",
`Honduran`=	"B03001_011",
`Nicaraguan`=	"B03001_012",
`Panamanian`=	"B03001_013",
`Salvadoran`=	"B03001_014",
`Other Central American`=	"B03001_015")

hispanic_CA<-get_acs(state = "MA", geography = "tract", 
                  variables = hispanic_CA, year = 2019, geometry = TRUE, 
                  summary_var = "B03001_008")
hispanic_CA<-hispanic_CA%>% mutate(Percentage = 100 * (estimate / summary_est))
head(hispanic_CA)

pdf("/Users/carmenrodriguez/Desktop/Spring2022/BST300/Spring 2022/R files/Figures/hispanic_CA.pdf")
hispanic_CA %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Hispanic or Latino: Central American groups by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()
dev.off()

#South America
hispanic_SA<-c(`Argentinean`=	"B03001_017",
`Bolivian`=	"B03001_018",
`Chilean`	="B03001_019",
`Colombian`=	"B03001_020",
`Ecuadorian`=	"B03001_021",
`Paraguayan`=	"B03001_022",
`Peruvian`=	"B03001_023",
`Uruguayan`	= "B03001_024",
`Venezuelan`=	"B03001_025",
`Other South American`=	"B03001_026")

hispanic_SA<-get_acs(state = "MA", geography = "tract", 
                  variables = hispanic_SA, year = 2019, geometry = TRUE, 
                  summary_var = "B03001_016")
hispanic_SA<-hispanic_SA%>% mutate(Percentage = 100 * (estimate / summary_est))
head(hispanic_SA)

pdf("/Users/carmenrodriguez/Desktop/Spring2022/BST300/Spring 2022/R files/Figures/hispanic_SA.pdf")
hispanic_SA %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Hispanic or Latino: South American groups by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()
dev.off()
```


```{r eval=FALSE, include=FALSE}
#pdf("/Users/carmenrodriguez/Desktop/Spring2022/BST300/Spring 2022/R files/Figures/HispanicHeritage.pdf")
hispanic_overall %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = " Hispanic or Latino: All groups by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

hispanic_CA %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Hispanic or Latino: Central American groups by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

hispanic_SA %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Hispanic or Latino: South American groups by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()
dev.off()
```

###---- Asian 
```{r eval=FALSE, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}

##- Divided by regions
#-- East Asian: China, Korea, Taiwan, Japan, Macau, Mongolia
asia_east<-c(`Chinese`=	"B02015_007",
`Japanese`=	"B02015_011",
`Okinawan`=	"B02015_017",
`Korean`=	"B02015_012",
`Mongolian`=	"B02015_015",
`Taiwanese`	="B02015_020",
`Hmong`	="B02015_009")
#Hmong mainly live in southwest china

#-- South Asia: Sri Lanka, Bangladesh, India, Afghanistan, Pakistan, Bhutan, Nepal, Iran, and the Maldives.
asia_south<-c(`Asian Indian`="B02015_002",
`Bangladeshi`="B02015_003",
`Bhutanse`=	"B02015_004",
`Nepalese`=	"B02015_016",
`Pakistani`	="B02015_018",
`Sri Lankan`=	"B02015_019")
#Burmese - Bangladeshi-- interesting because this is a language

#--Southeast Asia: Brunei, Cambodia, Indonesia, Laos, Malaysia, Myanmar, Philippines, Singapore, Thailand, Timor Leste, and Vietnam.
asia_southeast<-c(`Burmese`	="B02015_005",
  `Cambodian`="B02015_006",
  `Filipino`=	"B02015_008",
  `Indonesian`=	"B02015_010",
  `Malaysian`=	"B02015_014",
  `Laotian` ="B02015_013",
  `Thai`=	"B02015_021",
  `Vietnamese`=	"B02015_022")

asia_other<- c(`Other(specified)`	="B02015_023",
  `Other(not specified)`=	"B02015_024",
  `Two or more`	="B02015_025")

options(tigris_use_cache = TRUE) # cache shapefiles for use in future sessions

asia_east<-get_acs(state = "MA", geography = "tract", 
                  variables = asia_east, year = 2019, geometry = TRUE, 
                  summary_var = "B02015_001")

asia_south<-get_acs(state = "MA", geography = "tract", 
                  variables = asia_south, year = 2019, geometry = TRUE, 
                  summary_var = "B02015_001")

asia_southeast<-get_acs(state = "MA", geography = "tract", 
                  variables = asia_southeast, year = 2019, geometry = TRUE, 
                  summary_var = "B02015_001")

asia_other<-get_acs(state = "MA", geography = "tract", 
                  variables = asia_other, year = 2019, geometry = TRUE, 
                  summary_var = "B02015_001")

asia_east<-asia_east %>% mutate(Percentage = 100 * (estimate / summary_est))
asia_south<-asia_south %>% mutate(Percentage = 100 * (estimate / summary_est))
asia_southeast<-asia_southeast %>% mutate(Percentage = 100 * (estimate / summary_est))
asia_other<-asia_other %>% mutate(Percentage = 100 * (estimate / summary_est))


pdf("/Users/carmenrodriguez/Desktop/Spring2022/BST300/Spring 2022/R files/Figures/AsianAlone.pdf")
asia_east %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Asian Alone: East Asian by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

asia_south %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Asian Alone: South Asian by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

asia_southeast %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Asian Alone: South East Asian by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

asia_other %>% ggplot(aes(fill = Percentage)) + 
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Asian Alone: Other Asian by Census Tracts in MA",
       caption = "Data source: 2019 5-year ACS",
       fill = "Percent") + theme_minimal()

dev.off()
```



```{r,cache =TRUE}
options(tigris_use_cache = TRUE)
#2006-2010
vars<-c(`High school graduate or higher` = "DP02_0066P",
                 `Bachelors degree or higher` = "DP02_0067P",
                 `Below Poverty Line` = "DP03_0119P")
  
options(tigris_use_cache = TRUE) 
demographics_2010<-get_acs(state = "MA", geography = "tract", 
                  variables = vars, year = 2010, geometry = TRUE)  


#quick extraction to make the optimal care map
acstigris<-get_acs(state = "MA", geography = "tract", 
                  variables = "DP02_0066P", year = 2019, geometry = TRUE)


# demographics_2010 %>% filter(variable== "Median Income") %>% ggplot(aes(fill = estimate)) + 
#   geom_sf(color = NA) + 
#   coord_sf(crs = 6491, datum = NA) +
#   scale_fill_viridis_c(option = "C",direction = 1) +
#   labs(title = "Median Income Distribution by Census Tracts in MA",
#        caption = "Data source: 2006-2010 5-year ACS",
#        fill = "Estimate") + theme_minimal()

demographics_2010 %>% filter(variable %in% c("High school graduate or higher","Bachelors degree or higher")) %>% ggplot(aes(fill = estimate)) + 
  facet_wrap(~ variable) +                            
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Educational Attainment by Census Tracts in MA",
       caption = "Data source: 2006-2010 5-year ACS",
       fill = "Percentage") + theme_minimal()

demographics_2010 %>% filter(variable==c("Below Poverty Line")) %>% ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Families and people whose income is below poverty level by Census Tracts in MA",
       caption = "Data source: 2006-2010 5-year ACS",
       fill = "Percentage") + theme_minimal()
```

```{r,cache =TRUE}
#2011-2015
options(tigris_use_cache = TRUE)
demographics_2015<-get_acs(state = "MA", geography = "tract", 
                  variables = vars, year = 2015, geometry = TRUE) 


# demographics_2015 %>% filter(variable== "Median Income") %>% ggplot(aes(fill = estimate)) + 
#   geom_sf(color = NA) + 
#   coord_sf(crs = 6491, datum = NA) +
#   scale_fill_viridis_c(option = "C",direction = 1) +
#   labs(title = "Median Income Distribution by Census Tracts in MA",
#        caption = "Data source: 2011-2015 5-year ACS",
#        fill = "Estimate") + theme_minimal()


demographics_2015 %>% filter(variable %in% c("High school graduate or higher","Bachelors degree or higher")) %>% ggplot(aes(fill = estimate)) + 
  facet_wrap(~ variable) +                            
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Educational Attainment by Census Tracts in MA",
       caption = "Data source: 2011-2015 5-year ACS",
       fill = "Percentage") + theme_minimal()

demographics_2015 %>% filter(variable==c("Below Poverty Line")) %>% ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Families and people whose income is below poverty level by Census Tracts in MA",
       caption = "Data source: 2011-2015 5-year ACS",
       fill = "Percentage") + theme_minimal()


```

```{r,cache =TRUE}
#2015-2019
vars1<-c(`High school graduate or higher` = "DP02_0067P",
                 `Bachelors degree or higher` = "DP02_0068P",
                 `Below Poverty Line` = "DP03_0119P",
                  `Renter Occupied Housing` = "DP04_0047P",
                   `Owner Occupied Housing` = "DP04_0046P")
  
options(tigris_use_cache = TRUE) 
demographics_2019<-get_acs(state = "MA", geography = "tract", 
                  variables = vars1, year = 2019, geometry = TRUE)

# demographics_2019 %>% filter(variable== "Median Income") %>% ggplot(aes(fill = estimate)) + 
#   geom_sf(color = NA) + 
#   coord_sf(crs = 6491, datum = NA) +
#   scale_fill_viridis_c(option = "C",direction = 1) +
#   labs(title = "Median Income Distribution by Census Tracts in MA",
#        caption = "Data source: 2015-2019 5-year ACS",
#        fill = "Estimate") + theme_minimal()

demographics_2019 %>% filter(variable %in% c("High school graduate or higher","Bachelors degree or higher")) %>% ggplot(aes(fill = estimate)) + 
  facet_wrap(~ variable) +                            
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Educational Attainment by Census Tracts in MA",
       caption = "Data source: 2015-2019 5-year ACS",
       fill = "Percentage") + theme_minimal()

demographics_2019 %>% filter(variable==c("Below Poverty Line")) %>% ggplot(aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Families and people whose income is below poverty level by Census Tracts in MA",
       caption = "Data source: 2015-2019 5-year ACS",
       fill = "Percentage") + theme_minimal()

demographics_2019 %>% filter(variable %in% c("Renter Occupied Housing","Owner Occupied Housing")) %>% 
  ggplot(aes(fill = estimate)) +  
  facet_wrap(~ variable) +
  geom_sf(color = NA) + 
  coord_sf(crs = 6491, datum = NA) +
  scale_fill_viridis_c(option = "C",direction = 1) +
  labs(title = "Housing tenure by Census Tracts in MA",
       caption = "Data source: 2015-2019 5-year ACS",
       fill = "Percentage") + theme_minimal()

```


